\newpage
\chapter{Searching for dark matter by identifying invisible (or semi-visible) jets (13/12/2016)}
\label{sec:svjoverview}

One method of detecting dark matter -- that I'll probably be looking at in the foreseeable future -- is from the production of invisible (or semi-visible) jets. If dark matter is a composite particle that can couple to Standard Model particles via a weak portal, these "hidden-sector" states could be produced (and then undergo a QCD-like shower) at the LHC. So a jet/spray of \emph{stable} dark matter particles would be produced along with \emph{unstable} states that decay into Standard Model particles. So if the production and decay of these particles are driven by a $Z'$ (dark matter--Standard Model mediator) resonance that's leptophobic (doesn't couple strongly to leptons), the signature that you'd detect would be significant MET aligned roughly in the direction of a hadron jet. This is all explained well in \cite{Cohen:2015toa}.

These channels haven't been probed very well at the LHC so far because of the types of searches that have been, and currently are, conducted. The minimum azimuthal separation (difference in azimuthal angle) between jets is

\begin{equation}
\Delta\phi \equiv \min \{ \Delta\phi_{j_1 \ \etmiss}, \Delta\phi_{j_2 \ \etmiss} \}
\end{equation}

where $j_{1, 2}$ are the two hardest (highest momentum) jets \cite{Cohen:2015toa}. Typical searches require $\Delta\phi \gtrsim 0.4$, which is quite a bit larger than what are expected in the semi-visible jet channels. When including other cuts on, i.e., \etmiss and \alphat, the efficiency in this type of search drops to low values such that little data can be collected for interpretation. The horseshoe symbol $\supset$ indicates a superset. So the left-hand side of the formula is a superset of the right-hand side, i.e., the left-hand side includes, but is not limited to, the contents of the right-hand side. In the case of the Lagrangian, only the dark sector particles we are interested in are written, and it is implied that the full Lagrangian contains that, as well as more terms representing the other particles and interactions.

In this scenario, the dark matter is strongly coupled (i.e., couples to the strong force). One of the leading dark matter candidates, the WIMP, is weakly coupled. The paper compares sensitivity, \etmiss\ and $\Delta\phi$ between the two.

% CONTINUE WITH THIS SECTION WHEN I START DOING MORE WORK ON THIS