\newpage
\chapter{Introduction, software/hardware needs, stuff to remember (26/09/2016)}

\begin{easylist}[itemize]
\ListProperties(Style*=-- , FinalMark={)})
& Generally, be in the physics building 9--5 on weekdays.
& Bank holidays off.
& $\sim$ 4 weeks of leave.

& Should be getting my Macbook Air soon, then I can get up and running with proper work.
& In the meantime, install a Linux virtual machine (using VirtualBox) so I can use the computing software with my Fujitsu laptop.
\end{easylist}

Software to install on Linux VM/Mac:

\begin{easylist}[itemize]
\ListProperties(Style*=, FinalMark={)})
& \ROOT
& \madgraph
& \PYTHIA
\end{easylist}

For postgrad lecture courses I'll need to enroll in, see \url{http://mpags.nottingham.ac.uk/mpags} (username: eshwen, password: $<$normal password$>$). I need to enroll in 4 modules this term and 2 next term $\rightarrow$ statistics, C++, particle physics modules, quantum field theory?

\begin{easylist}[itemize]
\ListProperties(Style*=-- , FinalMark={)})
& Need to do some service work (Level 1 Trigger, most likely) and contribute so I can make the authors list.
& Will need to be proficient in C++ (so re-familiarise myself with C) and Python (scripting) because these are the languages used in the software I'll be using (\ROOT, etc.).
& Need to get registered with CERN.
- Will normally be in room 4.46 (postgrad office). Access code = 412.
\end{easylist}

So I've installed Linux Mint 18 (Cinammon) as a virtual machine on my Fujitsu after numerous failed attempts with other operating systems. But now it won't connect to the Internet to fetch updates for drivers and codecs, etc. And without internet, I can't install the necessary programs. \underline{So get the internet connection sorted!}

\textasciicircum\textasciicircum\textasciicircum Not worth the hassle of setting up a VM. I've just installed Cygwin for the time being, so I can use \PYTHIA and \ROOT, etc.

For the postgrad lecture courses, it's recommended that I take:
\begin{easylist}[itemize]
\ListProperties(Style*=-- , FinalMark={)})
& PP1: Introduction to Particle Physics
& PP3: QED and the Standard Model
& PP5: Group Theory (for Particle Physics)
& PP6: Particle Physics Techniques
\end{easylist}

The times of the lectures are on the webpage (see page 2/3), along with the lecture notes and other material.

The lectures are livestreamed in the lecture room 3.29 (opposite Martisse's office).

Statistics course (also recommended I take) in term 2. Takes place 9\textsuperscript{th}--13\textsuperscript{th} Jan. 2017, 11am--1pm in room 3.29.

For teaching/demonstrating, I'm doing first year maths tutorials in term 2 (weeks 13--24). Should be paid \pounds12.22 per hour. Contract should be issued around Christmas; will need my National Insurance number and passport.

Remember I also have to attend lectures for PHYSM3407 Current Topics in Physics (starts Tuesday 1\textsuperscript{st} November), focusing on the particle physics strand. I won't have to do any assessments, just attend lectures. It runs from weeks 6--11.

Guidelines for CMS publications (font, writing style, spelling, and any other subtleties) are clarified at \url{https://twiki.cern.ch/twiki/bin/viewauth/CMS/Internal/PubGuidelines}.
